---
title: "R Notebook"
output: html_notebook
---

```{r}
library(phyloseq); packageVersion("phyloseq")
library(tidyverse); packageVersion("tidyverse")
library(philr); packageVersion("philr")
library(ape); packageVersion("ape")

data(GlobalPatterns)

```


# From Phylr Docuumentation

## Filter Low Abundance OTUs

```{r}
GP <-  filter_taxa(GlobalPatterns, function(x) sum(x > 3) > (0.2*length(x)), TRUE)
GP <-  filter_taxa(GP, function(x) sd(x)/mean(x) > 3.0, TRUE)
GP <- transform_sample_counts(GP, function(x) x+1)
GP
```

## Process Phylogenetic Tree

```{r}
is.rooted(phy_tree(GP)) # Is the tree Rooted?
is.binary.tree(phy_tree(GP)) # All multichotomies resolved?
phy_tree(GP) <- makeNodeLabel(phy_tree(GP), method="number", prefix='n')
name.balance(phy_tree(GP), tax_table(GP), 'n1')
```

## Investigate Dataset Components

```{r}
otu.table <- t(otu_table(GP))
tree <- phy_tree(GP)
metadata <- sample_data(GP)
tax <- tax_table(GP)

otu.table[1:2,1:2] # OTU Table
```

```{r}
tree 
```

```{r}
head(metadata,2) # Metadata
```

```{r}
head(tax,2) # taxonomy table
```

## Transform data using PhILR

```{r}
gp.philr <- philr(otu.table, tree, 
                  part.weights='enorm.x.gm.counts', 
                  ilr.weights='blw.sqrt')
```

## Ordination
```{r}
gp.dist <- dist(gp.philr, method="euclidean")
gp.pcoa <- ordinate(GP, 'PCoA', distance=gp.dist)
plot_ordination(GP, gp.pcoa, color='SampleType') + geom_point(size=4)
```

# New Direciton
The authors suggest using glmnet to do a lasoo to identify whihc variables
best distinguis humans from non humans

```{r}
sample_data(GP)$human <- factor(get_variable(GP, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue"))
```

```{r}
as(sample_data(GP), "data.frame")%>% select(SampleType) -> sam
```

```{r}
merge(sam, gp.philr) -> gpsam
gpsam %>%
  mutate(SampleType = as.character(SampleType))  %>%
  mutate(BroadType = if_else(SampleType %in% c("Feces", "Mock", "Skin", "Tongue"), "Human", 
                             if_else(SampleType %in% c("Freshwater", "Ocean", "Freshwater (creek)"), "Water", "Other"))) %>%
  select(BroadType, everything()) %>%
  select(-SampleType) -> gpsam2
```

Now we have a data frame with all of the nodes, but also sample types. 

Lets just focus on the first 50 nodes, to keep this managable.

```{r}
gpsam2 <- gpsam2 %>% select(BroadType, n1:n50)
```

We want to know if BroadType predicts each node

```{r}
# Make long format
gpsamLong <- gpsam2 %>% pivot_longer(cols = n1:n50, names_to = "node", values_to = "ILR")
# 
gpsamNested <- gpsamLong %>% group_by(node) %>% nest()
```

Now we have a data frame, where one column is the nodes, and the other is a bunch of data frames for each node. Lets pull the data for the first node and model it.

```{r}
testNode1 <- gpsamNested %>% filter(node =="n1") %>% pull(data)
testNode1
```

```{r}
lm(ILR ~ BroadType, data = testNode1)
```

There was a problem with the merge, but this error especially still doesn't make sense.


